---
title: Cost-effectiveness analysis of public health interventions against RSV in Kenya
  and South Africa
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Loading required libraries

We need to load libraries required for the analysis.

```{r,results='hide',message=FALSE}

# html_document

rm(list=ls())
package_names <- c("tidyverse","rstudioapi","fitdistrplus","rstudioapi","matrixStats", # "ungeviz",
  "stringi","cowplot","here","conflicted","rriskDistributions","rgdal")
# if rriskDistributions does not install, try installing the following libraries from the command line:
# sudo apt install tk-dev tk-table

lapply(package_names, function(x) if (!any(row.names(installed.packages()) %in% x)) {install.packages(x)})
lapply(package_names,library,character.only=TRUE) # [!package_names %in% "here"]
currentdir_path=dirname(rstudioapi::getSourceEditorContext()$path); setwd(currentdir_path)
# library(here)
conflict_prefer("select", "dplyr"); conflict_prefer("filter", "dplyr")
num_sim <- 5000; source('functions/RSV_load_all.R')

```

### Loading functions

```{r,message=FALSE,results=FALSE}

source('functions/load_config_pars.R') # here::here(
lapply(c("functions/set_xlims_cea.R","functions/get_burden_flexible.R",
         "functions/get_burden_flexible_ari_sari.R",'functions/GammaParmsFromQuantiles.R',
         "functions/load_own_data.R"),function(x) {source(x)}) # here::here()

```


### Load incidence data from Kenya
```{r,message=FALSE,results=FALSE}

kenya_data_file_path <- "custom_input/Kenya_ARI_SARI_Rates_2010_2018_tidydata_updated_2021_08.csv"

### load Kenya ARI/SARI incidence data and structure as dataframe
kenya_ari_sari_incidence <- bind_rows(fcn_load_kenya(kenya_data_path=kenya_data_file_path,sel_disease="ARI")$rsv_incidence_ageinf,
          fcn_load_kenya(kenya_data_path=kenya_data_file_path,sel_disease="SARI")$rsv_incidence_ageinf) %>% 
  mutate(disease_type_medic_status=paste(disease_type, 
                                         ifelse(medically_attended,"medically attended","not attended")) ) %>%
  mutate(disease_type_medic_status=gsub("SARI medically attended","SARI hospitalised",
                                        disease_type_medic_status)) %>%
  mutate(disease_type_medic_status=gsub("SARI not attended","SARI non-hospitalised",
                                        disease_type_medic_status)) %>%
  dplyr::select(!c(RSV_assoc,freq,n)) %>% relocate(age_inf,.before=variable) %>% 
  relocate(metric_per_popul,.after=value) %>% 
  group_by(age_in_months,disease_type_medic_status) %>% summarise(value=unique(value),
    CI_95_lower=unique(CI_95_lower),CI_95_upper=unique(CI_95_upper),
    metric_per_popul=unique(metric_per_popul))

```

### PLOT Kenya incidence data

```{r,out.width='100%'}

popul_denom <- 1e5 # per person or per 100K?
# helper df to have axis limits fixed per row
df2 <- data.frame(disease_type_medic_status=unique(kenya_ari_sari_incidence$disease_type_medic_status),
                  age_in_months=1,value=c(rep(0.55,2),rep(0.125,2))*popul_denom)

ggplot(kenya_ari_sari_incidence, aes(x=age_in_months)) + 
  geom_bar(aes(y=popul_denom*value/metric_per_popul,fill=disease_type_medic_status),position="stack",stat="identity") +
  geom_errorbar(aes(ymin=popul_denom*CI_95_lower/metric_per_popul,ymax=popul_denom*CI_95_upper/metric_per_popul),size=0.4) +
  geom_point(data=df2,aes(x=age_in_months,y=value),color="white") +
  facet_wrap(~disease_type_medic_status,nrow=2,scales="free") + # 
  xlab("age (months)") + ylab(paste0("cases per ",ifelse(popul_denom>1,"100,000 ",""), "person year",ifelse(popul_denom>1,"s",""))) +
  labs(fill="")+ scale_y_continuous(expand=expansion(0.01,0)) +
  theme_bw() + standard_theme + theme(axis.text.x=element_text(vjust=0.5,size=8),axis.text.y=element_text(size=8),
               legend.text=element_text(size=7),legend.background=element_rect(fill=NA),legend.position="bottom", # c(0.92,0.925)
                                      axis.title.x=element_text(size=10),axis.title.y=element_text(size=10),
                                      strip.text=element_text(size=11))

save_flag=FALSE
if (save_flag){
ggsave(paste0("output/cea_plots/kenya_ari_sari_burden_errorbars_grouped_",ifelse(popul_denom>1,"per100k",""),".png"),
        width=42,height=22,units="cm") }

```

### Fit the uncertainty in ARI/SARI incidence data for Kenya with gamma distributions

```{r,message=FALSE,results=FALSE}

ci50_range <- c(25,75)/1e2; ci95_range <- c(2.5,97.5)/1e2
kenya_nonhosp_hosp_incid_ari_sari <- lapply(c("ARI","SARI"), function(x)
  fcn_gen_nonhosp_hosp_incid_samples_kenya(kenya_data_file_path,sel_disease=x,n_iter=5e3,age_maxval=60,
            CI_intervals=ci95_range,randsampl_distrib_type="gamma"))
names(kenya_nonhosp_hosp_incid_ari_sari)=c("ARI","SARI")

```

### Load Kenya deaths data & fit CI95 values with gamma distributions again

```{r,message=FALSE}

# Load data
deaths_kenya <- read_csv("custom_input/deaths_kenya_tidy_adjusted_02_2022.csv") %>% 
  filter(variable=="rate" & !age_in_months %in% c("<12","12-23","<24","24-59","<60")) %>% 
  mutate(age_in_months=ifelse(age_in_months=="<1","0",age_in_months),freq=1) %>% 
  mutate(freq=ifelse(grepl('-',age_in_months),as.numeric(sapply(age_in_months, 
                              function(x) diff(as.numeric(unlist(strsplit(x,"-"))))))+1,freq)) %>% 
  mutate(age_in_months_orig=factor(age_in_months,levels=unique(age_in_months)),
         age_in_months=ifelse(grepl('-',age_in_months), 
                              sapply(strsplit(age_in_months,'-'),'[[',1),age_in_months)) %>%
  uncount(weights=freq, .id="n",.remove=F) %>% mutate(age_inf=as.numeric(age_in_months)+(n-1)) %>% 
  dplyr::select(!c(n,freq,age_in_months)) %>% relocate(age_inf,.before=value)

# Fit
kenya_deaths_distrib_params <- bind_rows(lapply(c("yes","no"), function(y_no) data.frame(age_inf=0:59, 
  t(sapply(1:(nrow(deaths_kenya)/2), function(x) gamma.parms.from.quantiles(p=c(2.5,97.5)/100, 
  q=as.numeric((deaths_kenya %>% mutate(CI_95_lower=ifelse(CI_95_lower==0,0.1,CI_95_lower)) %>% 
                  filter(in_hospital==y_no) %>% 
  dplyr::select(c(CI_95_lower,CI_95_upper)))[x,]))[c("shape","rate")])),in_hospital=y_no))) %>% 
  mutate(shape=unlist(shape),rate=unlist(rate))
```

### Generate samples for Kenya deaths from the fitted distributions

```{r}

kenya_deaths_incid <- lapply(c("yes","no"), function(y_no) t(sapply(0:59, function(x) 
    rgamma(5e3,shape=(kenya_deaths_distrib_params %>% filter(age_inf==x&in_hospital==y_no))$shape,
          rate=(kenya_deaths_distrib_params %>% filter(age_inf==x&in_hospital==y_no))$rate)))/1e5)
names(kenya_deaths_incid)=c("hosp","nonhosp")

```

### Load deaths data for South Africa

```{r}
deaths_SA <- read_csv("custom_input/mortality South Africa updated17_01_2022_TIDY.csv") %>% 
  rename(age_in_months=`age in months`,CI_95_lower=LCI,CI_95_upper=UCI,in_hospital=`medically attended`) %>% 
  filter(!in_hospital %in% "total") %>%
  mutate(in_hospital=ifelse(grepl("in-hospital",in_hospital),"yes","no")) %>%
  filter(variable %in% "rate" & !age_in_months %in% c("<1 year","<5 years")) %>% 
  mutate(age_in_months=ifelse(age_in_months=="<1","0",age_in_months),freq=1) %>% 
  mutate(freq=ifelse(grepl('-',age_in_months),as.numeric(sapply(age_in_months, 
                                      function(x) diff(as.numeric(unlist(strsplit(x,"-"))))))+1,freq)) %>% # 
  mutate(age_in_months_orig=factor(age_in_months,unique(age_in_months)),
         age_in_months=ifelse(grepl('-',age_in_months), 
                              sapply(strsplit(age_in_months,'-'),'[[',1),age_in_months)) %>%
  uncount(weights=freq, .id="n",.remove=F) %>% mutate(age_inf=as.numeric(age_in_months)+(n-1)) %>% 
  dplyr::select(!c(n,freq,age_in_months)) %>% relocate(age_inf,.before=value)

# plot separately
# ggplot(deaths_SA,aes(x=age_inf,y=value,color=in_hospital)) + geom_line() + geom_point() +
#   geom_ribbon(aes(ymin=CI_95_lower,ymax=CI_95_upper,fill=in_hospital),alpha=0.2) + theme_bw() +
#   scale_x_continuous(breaks=(0:30)*2,expand=expansion(0.01,0)) + scale_y_log10(breaks=2^(-3:7))
```

### Plot incidence of deaths/100k population for both countries

```{r,out.width='100%'}

deaths_data <- bind_rows(deaths_SA %>% select(!age_inf) %>% group_by(age_in_months_orig,in_hospital) %>%
            summarise(value=unique(value),CI_95_lower=unique(CI_95_lower),CI_95_upper=unique(CI_95_upper),
            country="South Africa") %>% ungroup() %>% 
            mutate(age_in_months_orig=as.character(age_in_months_orig)),
        deaths_kenya %>% group_by(age_in_months_orig,in_hospital) %>% summarise(value=unique(value),
            CI_95_lower=unique(CI_95_lower),CI_95_upper=unique(CI_95_upper),country="Kenya")) %>%
        mutate(age_in_months_orig=factor(age_in_months_orig,levels=unique(age_in_months_orig))) %>% 
        group_by(country,age_in_months_orig) %>% 
        mutate(CI_95_lower_sum=sum(CI_95_lower),CI_95_upper_sum=sum(CI_95_upper)) %>% 
  mutate(in_hospital=ifelse(in_hospital=="yes","in-hospital","out-of-hospital"))

# PLOT
ggplot(deaths_data,aes(x=age_in_months_orig)) + 
  geom_bar(aes(y=value,fill=in_hospital),stat="identity") + # ,position="dodge"
  geom_errorbar(aes(ymin=CI_95_lower_sum,ymax=CI_95_upper_sum,group=in_hospital),size=0.2) + # ,position="dodge"
  facet_wrap(~country) + # ,nrow=2 # ,scales="free_y"
  xlab("age (months)") + ylab("deaths per 100,000 person year") + labs(fill="") +
  scale_y_continuous(expand=expansion(0.01,0),breaks=(0:12)*25) + theme_bw() + standard_theme + 
  theme(axis.text.x=element_text(vjust=0.5,size=11),axis.text.y=element_text(size=11),
    legend.text=element_text(size=13),legend.background=element_rect(fill=NA),strip.text=element_text(size=14),
    legend.position=c(0.88,0.925),axis.title.x=element_text(size=13),axis.title.y=element_text(size=13))

# SAVE
# ggsave("output/cea_plots/ALL_deaths_data_stacked_yfixed.png",width=32,height=18,units="cm")

# Calculate concentration of deaths in first 5 months
# (this is a bit crude bc assuming age group size = duration)
deaths_data %>% mutate(before_5_mts=as.numeric(age_in_months_orig)<=5,
                       size_agegr=ifelse(as.numeric(age_in_months_orig)<=12,1,
                                         ifelse(as.numeric(age_in_months_orig)<=16,2,12))) %>%
   group_by(country,before_5_mts) %>%  summarise(value=sum(value*size_agegr))

```

### Fit CI95s of South African deaths data with gamma distributions

```{r}
SA_deaths_distrib_params <- bind_rows(lapply(c("yes","no"), 
        function(y_no) data.frame(age_inf=0:(nrow(deaths_SA)/2-1), 
        t(sapply(1:(nrow(deaths_SA)/2), function(x) gamma.parms.from.quantiles(p=c(2.5,97.5)/100, 
        q=as.numeric((deaths_SA %>% mutate(CI_95_lower=ifelse(CI_95_lower==0,0.1,CI_95_lower)) %>% 
        filter(in_hospital==y_no) %>% 
        dplyr::select(c(CI_95_lower,CI_95_upper)))[x,]))[c("shape","rate")])),in_hospital=y_no))) %>% 
  mutate(shape=unlist(shape),rate=unlist(rate))
# generate samples from fitted distribs
popul_denom<-unique(deaths_SA$metric_per_popul)
SA_deaths_incid <- lapply(c("yes","no"), function(y_no) t(sapply(0:(nrow(deaths_SA)/2-1), function(x) 
  rgamma(5e3,shape=(SA_deaths_distrib_params %>% filter(age_inf==x&in_hospital==y_no))$shape,
         rate=(SA_deaths_distrib_params %>% filter(age_inf==x&in_hospital==y_no))$rate)))/popul_denom)
names(SA_deaths_incid)=c("hosp","nonhosp")

```

### Check if the fits match the data well

```{r,out.width='100%'}
ggplot(bind_rows(data.frame(age_inf=0:59,value=rowMeans(SA_deaths_incid$hosp),
            t(sapply(1:nrow(SA_deaths_incid$hosp),
    function(x) quantile(SA_deaths_incid$hosp[x,],probs=c(2.5,97.5)/100))),source="fit" ) %>%
      rename(CI_95_lower=`X2.5.`,CI_95_upper=`X97.5.`),
  deaths_SA %>% filter(in_hospital %in% "yes") %>%
      select(age_inf,value,CI_95_lower,CI_95_upper) %>%
    mutate(source="data",value=value/1e5,CI_95_lower=CI_95_lower/1e5,CI_95_upper=CI_95_upper/1e5) ),
      aes(x=age_inf,color=source,fill=source)) + geom_point(aes(y=value)) + geom_line(aes(y=value)) +
  geom_ribbon(aes(ymin=CI_95_lower,ymax=CI_95_upper),alpha=1/5) + scale_y_log10() +
  theme_bw() + standard_theme
```

### Load South African incidence data of RSV disease (ARI/SARI)

#### Load SARI data

```{r}
SA_SARI_data <- fcn_load_s_afr(safr_data_path = "custom_input/s_afr_incidence_data_rate.csv") %>%
  mutate(disease_type_medic_status=paste(disease_type,
        ifelse(hospitalisation,"hospitalised","not hospitalised")) ) %>%
  rename(agegroup_mts=age) %>% 
  relocate(age_inf,.before=Province) %>% relocate(Province,.after=disease_type_medic_status) %>% 
  relocate(year,.before=Province)
```

#### Load ILI (influenza-like illness) data
```{r}
SA_ILI_data <- read_csv("custom_input/s_afr_ILI_incidence_rate_160921.csv") %>% 
  filter(!(grepl("<",agegroup) | agegroup %in% c("0-5m","6-11m","12-23m","24-59m","<5y"))) %>%
  mutate(agegroup_mts=agegroup,agegroup=gsub("m","",agegroup), freq=1) %>% 
  mutate(freq=ifelse(grepl('-',agegroup),as.numeric(sapply(agegroup,
                function(x) diff(as.numeric(unlist(strsplit(x,"-"))))))+1,freq)) %>% 
  mutate(agegroup=ifelse(grepl('-',agegroup), sapply(strsplit(agegroup,'-'),'[[',1),agegroup)) %>%
  uncount(weights=freq, .id="n",.remove=F) %>% 
  mutate(age_inf=as.numeric(agegroup)+(n-1)) %>% dplyr::select(!c(n,freq,agegroup)) %>% # 
  relocate(age_inf,.before=rate) %>% relocate(disease_type,.before=hospitalisation) %>%
  mutate(disease_type=ifelse(disease_type=="ILI","ARI",""),
         disease_type_medic_status=ifelse(hospitalisation,paste0("medically attended ",disease_type),
                paste0("non medically attended ",disease_type))) %>% 
  relocate(disease_type_medic_status,.before=Province)
```

#### Concatenate two datasets
```{r}
if (!exists("SA_data")){ SA_data=bind_rows(SA_ILI_data,SA_SARI_data) }
```

### Plot South African disease incidence (ARI/SARI)

```{r,out.width='100%'}

# put it into one dataframe
SA_ILI_SARI_rawdata <- bind_rows( 
  distinct(fcn_load_s_afr(safr_data_path="custom_input/s_afr_incidence_data_rate.csv") %>%
  mutate(disease_type_medic_status=paste(disease_type,
                     ifelse(hospitalisation,"hospitalised","not hospitalised"))) %>% select(!age_inf)),
read_csv("custom_input/s_afr_ILI_incidence_rate_160921.csv") %>% 
  filter(!(grepl("<",agegroup) | agegroup %in% c("0-5m","6-11m","12-23m","24-59m","<5y"))) %>%
  rename(age=agegroup) %>% mutate(disease_type=ifelse(disease_type=="ILI","ARI",""),
         disease_type_medic_status=ifelse(hospitalisation,paste0("medically attended ",disease_type),
                                          paste0("non medically attended ",disease_type))) ) %>%
  mutate(age=gsub("m","",age),age=factor(age,levels=unique(age)))

# PLOT
plot_popul_denom <- 1e5 # per person or per 100K?
  # helper df to have axis limits fixed per row
  df2 <- data.frame(disease_type_medic_status=unique(SA_ILI_SARI_rawdata$disease_type_medic_status),age=1/2,
                    rate=c(rep(0.115,2),rep(0.32,2))*plot_popul_denom)
# create plot
ggplot(SA_ILI_SARI_rawdata,aes(x=age)) + 
  geom_bar(aes(y=plot_popul_denom*rate/popul_denom,fill=disease_type_medic_status),position="stack",stat="identity") +
  geom_errorbar(aes(ymin=plot_popul_denom*rate_CI_lower/popul_denom,ymax=plot_popul_denom*rate_CI_upper/popul_denom),size=0.4) +
  geom_point(data=df2,aes(x=age,y=rate),color="white") +
  facet_wrap(~disease_type_medic_status,nrow=2,scales = "free") + # 
  xlab("age (months)") + ylab(paste0("cases per ",ifelse(popul_denom>1,"100,000 ",""), "person year",ifelse(popul_denom>1,"s",""))) +
  labs(fill="") + scale_y_continuous(expand=expansion(0.01,0)) +
  theme_bw() + standard_theme + theme(axis.text.x=element_text(vjust=0.5,size=8),axis.text.y=element_text(size=8),
               legend.text=element_text(size=7),legend.background=element_rect(fill=NA),legend.position="bottom", # c(0.92,0.925)
                                      axis.title.x=element_text(size=10),axis.title.y=element_text(size=10),
                                      strip.text=element_text(size=11))
# SAVE
# ggsave(paste0("output/cea_plots/SA_ari_sari_burden_errorbars_grouped_ILI_160921",ifelse(popul_denom>1,"per100k",""),".png"),
#        width=42,height=22,units="cm")

```


### Expanding ILI rates to ARI by taking % of ARI with fever in Kenya

```{r}
ILI_adjust_SA=TRUE
if (ILI_adjust_SA) { # 
  print("divide by fever proportion")
  SA_data[SA_data$disease_type=="ARI",c("rate","rate_CI_lower","rate_CI_upper")]=
  SA_data[SA_data$disease_type=="ARI",c("rate","rate_CI_lower","rate_CI_upper")]/mean(c(0.333,0.205,0.16))
  divided_fever=TRUE 
  }
```

### Fit South African ARI/SARI incidence data by fitting CI95s with gamma distributions

```{r}

# CI lower limit should not be 0 -> setting it to 1 in those cases where it's 0
# since average value of mean rate is > 1000, this does not change results more than 0.1%
SA_data$rate_CI_lower[SA_data$age_inf==0 & SA_data$disease_type=="ARI"]=1

### generate 5e3 sample paths for CEA
sa_nonhosp_hosp_incid_ari_sari=lapply(c("ARI","SARI"), 
      function(x) fcn_gen_nonhosp_hosp_incid_samples_SA(SA_data,diseasetype=x,
      n_iter=5e3, age_maxval=60,CI_intervals=ci95_range,randsampl_distrib_type="gamma"))
names(sa_nonhosp_hosp_incid_ari_sari)=c("ARI","SARI")



```

## Cost Estimates

### South African cost estimates

#### Inflation data

```{r}
inflation_data <- (read_csv("custom_input/inflation_data.csv") %>% filter(grepl("South Africa",`Country Name`)))
inflation_rate <- 1+as.numeric(inflation_data[,(ncol(inflation_data)-5):ncol(inflation_data)])/100
# ratio of exchange rates 2014 vs 2021
exch_rate_adj <- 11.2/14.77
# adjustment of 2014 prices
hist_adj <- prod(inflation_rate)*exch_rate_adj
```

### Load South Africa outpatient cost 

```{r}
s_afr_outpatient_cost <- bind_rows(data.frame(age="0-59",mean=25,LCI=18.3,UCI=31.8,
                                          cost_type="healthcare",disease="outpatient",name="total"),
      read_csv("custom_input/s_afr_PDE_calcs.csv") %>% filter(grepl("outpatient",disease)) %>%  
  rename(mean=`Mean cost per illness episode (USD)`)) %>% 
  # adjustment for inflation and exch rate change
  mutate(mean=mean*hist_adj,LCI=LCI*hist_adj,UCI=UCI*hist_adj)
```

### Fitting cost estimates (CI95s) with gamma distributions

```{r}
if (!any(grepl("shape",colnames(s_afr_outpatient_cost)))) {
  # colbind to original dataframe
for (k_row in 1:nrow(s_afr_outpatient_cost)) {
    gamma_fit <- c(get.gamma.par(p=ci95_range,q=c(s_afr_outpatient_cost$LCI[k_row],s_afr_outpatient_cost$UCI[k_row]),
                             show.output=F,plot=F),scaling=1)
  # if fitting fails, this can be overcome by dividing the costs by 100 and then scaling the fit 100x afterwards
  if (any(is.na(gamma_fit))) { 
    gamma_fit <- c(get.gamma.par(p=ci95_range,
                      q=c(s_afr_outpatient_cost$LCI[k_row],s_afr_outpatient_cost$UCI[k_row])/100,
                                                show.output=F,plot=F),scaling=100) 
    }
  sim_gamma <- rgamma(n=1e4,shape=gamma_fit["shape"],rate=gamma_fit["rate"])*gamma_fit["scaling"]
  gamma_fit <- c(gamma_fit,sim_mean=mean(sim_gamma),
                     sim_ci95_low=as.numeric(quantile(sim_gamma,probs=ci95_range[1])),
                     sim_ci95_up=as.numeric(quantile(sim_gamma,probs=ci95_range[2])))
  if (k_row==1) { SA_outpatient_fit_pars <- gamma_fit } else {
    SA_outpatient_fit_pars <- bind_rows(SA_outpatient_fit_pars,gamma_fit)} 
}
# bind together
  s_afr_outpatient_cost <- bind_cols(s_afr_outpatient_cost,SA_outpatient_fit_pars) %>% 
  mutate(freq=ifelse(grepl('-',age),
                     as.numeric(sapply(age, function(x) diff(as.numeric(unlist(strsplit(x,"-"))))))+1,1)) %>%
  mutate(age=ifelse(grepl('-',age), sapply(strsplit(age,'-'),'[[',1),age)) %>%
  uncount(weights=freq, .id="n",.remove=F) %>% # dplyr::
  mutate(age=as.numeric(age)+(n-1)) %>% select(!c(n,freq)) 
}

# PLOT: compare with and data
ggplot(s_afr_outpatient_cost) + geom_point(aes(x=mean,y=sim_mean,color=name,fill=cost_type),shape=21,size=3) +
   theme_bw() + scale_x_log10() + scale_y_log10(expand=expansion(0.1,0))

```
### Load inpatient costs (South Africa)

```{r}
s_afr_inpatient_cost <- read_csv("custom_input/s_afr_PDE_calcs.csv") %>% 
  filter(!name %in% "PDE" & grepl("inpatient",disease)) %>% 
  rename(mean=`Mean cost per illness episode (USD)`) %>%
  mutate(freq=ifelse(grepl('-',age),
                     as.numeric(sapply(age, function(x) diff(as.numeric(unlist(strsplit(x,"-"))))))+1,1)) %>% 
  mutate(age=ifelse(grepl('-',age), sapply(strsplit(age,'-'),'[[',1),age)) %>% 
  uncount(weights=freq, .id="n",.remove=F) %>% # dplyr::
  mutate(age=as.numeric(age)+(n-1)) %>% select(!c(n,freq)) %>%
  mutate(mean=mean*hist_adj,LCI=LCI*hist_adj,UCI=UCI*hist_adj) # adjustment for inflation and exch rate change

```

### Fit inpatient cost estimates (South Africa) with gamma distributions

```{r,warning=FALSE}
if (!any(grepl("shape",colnames(s_afr_inpatient_cost)))){
  sa_costs_unique <- s_afr_inpatient_cost %>% select(c(name,mean,LCI,UCI,cost_type,disease)) %>% distinct()
  for (k_row in 1:nrow(sa_costs_unique)) {
    g_fit_data <- c(ifelse(sa_costs_unique$LCI[k_row]==0,0.01,sa_costs_unique$LCI[k_row]),
                    sa_costs_unique$UCI[k_row])
    fit_gamma <- c(get.gamma.par(q=g_fit_data,p=ci95_range,show.output=F,plot=F)[c("shape","rate")],scaling=1)
    if (any(is.na(fit_gamma))) {
      print("first fit unsuccesful, re-fit with rescaling")
      fit_gamma <- c(get.gamma.par(q=g_fit_data/100,p=ci95_range,show.output=F,plot=F)[c("shape","rate")],scaling=100)
      }
    sim_gam<-rgamma(n=1e4,rate=fit_gamma["rate"],shape=fit_gamma["shape"])*fit_gamma["scaling"]
    gamma_out<-c(n=k_row,fit_gamma,sim_mean=mean(sim_gam),
                 sim_ci95_low=as.numeric(quantile(sim_gam,probs=ci95_range[1])),
                 sim_ci95_up=as.numeric(quantile(sim_gam,probs=ci95_range[2])))
    if (k_row==1) { gamma_fits <- gamma_out  } else { gamma_fits <- bind_rows(gamma_fits,gamma_out)}
  }
  s_afr_inpatient_cost <- left_join(s_afr_inpatient_cost,
            left_join(sa_costs_unique %>% mutate(n=row_number()),gamma_fits,by="n") %>% select(!n),
            by=c("name","mean","LCI","UCI","cost_type","disease"))
}

list_SA_costs <- list("inpatient"=s_afr_inpatient_cost,"outpatient"=s_afr_outpatient_cost)

# cost data types:
# list_SA_costs$inpatient %>% select(c(name,disease,cost_type)) %>% distinct()
# list_SA_costs$outpatient %>% select(c(name,disease,cost_type)) %>% distinct()
```

#### Check if (South Afr) inpatient cost fits are good

```{r,out.width='100%'}

ggplot(s_afr_inpatient_cost,aes(x=age)) + geom_line(aes(y=mean)) + geom_point(aes(y=mean),shape=21) +
  geom_line(aes(y=sim_mean),color="red") + geom_point(aes(y=sim_mean),shape=21,color="red") +
  geom_ribbon(aes(ymin=LCI,ymax=UCI),alpha=0.2) + geom_ribbon(aes(ymin=sim_ci95_low,ymax=sim_ci95_up),fill="red",alpha=0.2) +
  facet_wrap(disease~name~cost_type,scales="free_y") + theme_bw() + standard_theme + ggtitle("black: data, red = fitted values")

```
## KENYA cost estimates

### Load Kenya cost data

```{r}

kenya_costs <- read_csv("custom_input/kenya_costing_tables_tidy.csv")
# using inpatient/outpatient ratio in South Africa OR study from Malawi, taken from:
# https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7864144/
in_outpatient_cost_ratio <- 45.37/9.26
# we assume that inpatient/outpatient cost ratio is same as in Malawi, and total cost
# (# inpatients)*outpatient_cost*SA_in_outpatient_ratio + (# outpatients)*outpatient_cost = total cost
# outpatient_cost = (total cost)/[(# inpatients)*SA_in_outpatient_ratio + (# outpatients)]
KEN_outpatient_cost <- kenya_costs$mean[kenya_costs$variable %in% "Total healthcare cost"]/
  (kenya_costs$mean[grepl("Total number of inpatients",kenya_costs$variable)]*in_outpatient_cost_ratio + 
     kenya_costs$mean[grepl("Total number of outpatients",kenya_costs$variable)])
# assemble list
ken_inpatient_rows <- kenya_costs %>% filter(grepl("Siaya",site) & grepl("Total patient",variable))

```

### Fit Kenya cost data with gamma distributions

```{r,warning=FALSE}
list_KEN_costs <- list(inpatient_household=bind_cols(
  kenya_costs %>% filter(grepl("Siaya",site) & grepl("Total patient",variable)), 
  t(sapply(1:nrow(ken_inpatient_rows), function(x)
  unlist(get.gamma.par(q=c(ken_inpatient_rows$ci95_low[x],ken_inpatient_rows$median[x],
            ken_inpatient_rows$ci95_up[x])/100,p=c(2.5,50,97.5)/100,plot=F)[c("shape","rate")]))),scaling=100),
  inpatient_healthcare_system=KEN_outpatient_cost*in_outpatient_cost_ratio,outpatient_cost=KEN_outpatient_cost)
```

## Efficacy figures

```{r}
lists_of_effics <- list(
  # published novavax data (MV), for mAB: https://www.nejm.org/doi/pdf/10.1056/NEJMoa2110275?articleTools=true)
  list(mat_vacc=list(sympt_disease=c(mean=0.394,CI95_low=0.053,CI95_high=0.612),
                                         hospit=c(mean=0.444,CI95_low=0.196,CI95_high=0.615),
                                         severe=c(mean=0.483,CI95_low=-8.2/100,CI95_high=0.753), # c(mean=0.483,-8.2/100,0.753)
                                         half_life=36.5/30,duration=3),
                           monocl_ab=list(sympt_disease=c(mean=0.772,CI95_low=0.587,CI95_high=0.875),
                                hospit=c(mean=0.611,CI95_low=-3.7/100,CI95_high=0.854),
                                half_life=59.3/30,duration=5)),
  # Phase II data for nirsevimab (mAb): sympt_disease=c(mean=0.701,CI95_low=0.523,CI95_high=0.812),
                                          # hospit=c(mean=0.784,CI95_low=0.519,CI95_high=0.903),
                                          # half_life=59.3/30,duration=5))
  # new data (for MV: Pfizer data from conference, for mAB: https://www.nejm.org/doi/pdf/10.1056/NEJMoa2110275?articleTools=true)
                    list(mat_vacc=list(sympt_disease=c(mean=0.847,CI95_low=0.216,CI95_high=0.976),
                                hospit=c(mean=0.847,CI95_low=0.216,CI95_high=0.976), # no sep data
                                severe=c(mean=0.915,CI95_low=-5.6/100,CI95_high=0.998),
                                half_life=36.5/30,duration=3),
                           monocl_ab=list(sympt_disease=c(mean=0.772,CI95_low=0.587,CI95_high=0.875),
                                hospit=c(mean=0.611,CI95_low=-3.7/100,CI95_high=0.854),
                                half_life=59.3/30,duration=5)),
  # hypothetical, 40% mean effic for everything
  list(mat_vacc=list(sympt_disease=c(mean=0.4,CI95_low=0.2,CI95_high=0.6),
                     hospit=c(mean=0.4,CI95_low=0.2,CI95_high=0.6), # no sep data
                     severe=c(mean=0.4,CI95_low=0.2,CI95_high=0.6),half_life=36.5/30,duration=3),
       monocl_ab=list(sympt_disease=c(mean=0.4,CI95_low=0.2,CI95_high=0.6),
                      hospit=c(mean=0.4,CI95_low=0.2,CI95_high=0.6),half_life=59.3/30,duration=5)),
  # hypothetical, 60% mean effic for everything
  list(mat_vacc=list(sympt_disease=c(mean=0.6,CI95_low=0.4,CI95_high=0.8),
                     hospit=c(mean=0.6,CI95_low=0.4,CI95_high=0.8), # no sep data
                     severe=c(mean=0.6,CI95_low=0.4,CI95_high=0.8),half_life=36.5/30,duration=3),
       monocl_ab=list(sympt_disease=c(mean=0.6,CI95_low=0.4,CI95_high=0.8),
                      hospit=c(mean=0.6,CI95_low=0.4,CI95_high=0.8),half_life=59.3/30,duration=5)),
  list(mat_vacc=list(sympt_disease=c(mean=0.8,CI95_low=0.6,CI95_high=1),
                     hospit=c(mean=0.8,CI95_low=0.6,CI95_high=1), # no sep data
                     severe=c(mean=0.8,CI95_low=0.6,CI95_high=1),half_life=36.5/30,duration=3),
       monocl_ab=list(sympt_disease=c(mean=0.8,CI95_low=0.6,CI95_high=1),
                      hospit=c(mean=0.8,CI95_low=0.6,CI95_high=1),half_life=59.3/30,duration=5))
  )

efficacy_figures <- lists_of_effics[[1]]

```

### Fit efficacy by distributions

```{r}
# fitting efficacy figures with a beta distribution
source("functions/fit_efficacy.R")
g(list_effic_betafit,allfits) %=% fcn_betafit_efficacy(effic_figs=efficacy_figures,
                                   scan_range_resol_nsample=c(min=-2,max=2,by=1/100,n_sample=2e4),
                                   optim_range_res=c(min=-1,max=2,by=0.04),optim_initguess=c(-0.05,1))
# fit as: beta_fit <- rbeta(n=1e4,shape1=alphaval,shape2=betaval)*scale_val + shift_val

```
#### Shift distributions to fit the mean better? (set to false by default)

```{r}
# if this is set to true, then the fit is shifted to be more aligned with the mean efficacy rather than
# the CI interval (beta distribution often cannot do both)
flag_adjust_to_mean <- FALSE
if (flag_adjust_to_mean){
mv_severe_err <- c(t(allfits %>% filter(interv=="mat_vacc" & disease=="severe") %>% 
                       select(c(data_mean,data_CI95_low,data_CI95_high)))) -
  c(t(allfits %>% filter(interv=="mat_vacc"&disease=="severe") %>% select(c(mean,CI95_low,CI95_high))))
# adjust
list_effic_betafit$mat_vacc$severe["shift_fit"] <- list_effic_betafit$mat_vacc$severe["shift_fit"] + mv_severe_err[1]
# try how mean changed
sim_beta <- (rbeta(n=1e4,shape1=list_effic_betafit$mat_vacc$severe["shape1"],
              shape2=list_effic_betafit$mat_vacc$severe["shape2"])*list_effic_betafit$mat_vacc$severe["scale_fit"])+
  list_effic_betafit$mat_vacc$severe["shift_fit"]
c(mean(sim_beta),quantile(sim_beta,probs=ci95_range))
}
```

### Dose prices

```{r}
pricelist=list("mat_vacc"=c(3),"mAb"=c(6))
# loop thru: cntrs * interventions
cntrs_cea=c("KEN","ZAF")
par_table <- expand_grid(n_cntr_output=1:length(cntrs_cea),n_interv=1:2); read_calc_flag=c("calc","read")[1]
kenya_deaths_input=TRUE; SA_deaths_input=TRUE
```

### Set if exponential waning of efficacy, create folder name

```{r}
# exponential waning model used for efficacy
exp_wane_val <- FALSE
# parameters for exponential waning model
g(list_exp_waning_param,df_exp_waning_param) %=% fcn_exp_waning_rate(efficacy_figures,n_row=60)
# distribution used to fit efficacy figures
effic_dist_fit <- "beta" # 
lower_cov=FALSE # ; lower_cov_val=0.7
subfolder_name <- paste0("new_price_efficacy_",ifelse(kenya_deaths_input,"KENdeaths",""),
                       ifelse(SA_deaths_input,"_SAdeaths",""),
        "_CIs_SA_ILI_",ifelse(ILI_adjust_SA,"broader","narrow"),ifelse(exp_wane_val,"_expwaning",""),
        # ifelse(FALSE,paste0("_coverage",lower_cov_val),""),
        ifelse(flag_adjust_to_mean,"_adj_mv_sev_eff_mean",""),
        ifelse(grepl("gamma",effic_dist_fit),"","_effic_betafit"),"/") 
# ifelse(flag_publ_effic,"","_interim")
```

### Set up initial conditions
```{r,message=FALSE}
### Before starting loop for the FIRST TIME need to create temp folder (afterwards can comment out this line)
source("init_cea_calc_parallel.R")
```

### Outputs of CEA to display
```{r}
# outputs to display
all_cols=c("non_hosp_cases","hosp_cases",
           "rsv_deaths","rsv_deaths_disc",
           "total_DALY","total_DALY_disc",
           "cost_rsv_hosp",
           "total_medical_cost","total_medical_cost_disc",
           "incremental_cost","incremental_cost_disc",
           "total_medical_cost_averted","total_medical_cost_disc_averted",
           "non_hosp_cases_averted","hosp_cases_averted",
           "rsv_deaths_averted", 
           "total_DALY_averted","total_DALY_disc_averted",
           "SARI_averted",
           "incremental_cost/DALY_averted",
           "total_YLD","total_YLD_disc",
           "total_YLL","total_YLL_disc",
           "total_YLD_disc_averted","total_YLL_disc_averted",
           "hosp_SARI","non_hosp_SARI","ARI_averted") 
####
burden_cols <- all_cols[!grepl("cost|averted",all_cols)]; cost_cols <- all_cols[grepl("cost",all_cols)]
```


## RUN cost effectiveness calculations
```{r,message=F,results=F,warning=F}

ci50_range <- c(25,75)/1e2; ci95_range <- c(2.5,97.5)/1e2
# do we also want to calculate with projections from [Li 2020]
CALC_PROJECTION=TRUE
# run calculations (approx 3 mins if CALC_PROJECTION=FALSE, 6 mins if TRUE)
source("functions/cea_loop_cntr_interv.R")

```
